---
title: "Seurat_human_fetal_gut"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

rmarkdown::render(input = paste0("/nfs/team205/vk7/sanger_projects/cell2location_paper/notebooks/benchmarking/other_methods/Seurat_human_adult_oxford_gut.Rmd"), 
                  output_format = "html_document", 
                  output_file=paste0("/nfs/team205/vk7/sanger_projects/cell2location_paper/notebooks/benchmarking/other_methods/Seurat_human_adult_oxford_gut.html"))

## Seurat V3

```{r }
library(Matrix)
library(data.table)
library(Seurat)
library(SeuratDisk)
```

```{r}
# read cell2location model results 
sc_results_folder = '/nfs/team205/vk7/sanger_projects/collaborations/fetal_gut_mapping/results/'
results_folder = '/nfs/team205/vk7/sanger_projects/collaborations/fetal_gut_mapping/results/oxford/'

reg_mod_name = 'RegressionGeneBackgroundCoverageGeneTechnologyTorch_88covariates_124049cells_13904genes_noLN_revise'
reg_path = paste0(sc_results_folder, '', reg_mod_name, '/')
#Convert(paste0(reg_path,'sc_for_seurat.h5ad'), dest = "h5seurat", overwrite = TRUE)

sc_seu = LoadH5Seurat(paste0(reg_path,'sc_for_seurat.h5seurat'))
Idents(sc_seu) = sc_seu$combined
```

```{r}
sp_data_path = '/nfs/team205/vk7/sanger_projects/large_data/gut_kj_re/oxford_visium/'

Convert(paste0(sp_data_path, 'sp_for_seurat_selected.h5ad'), dest = "h5seurat", overwrite = TRUE)
sp_seu = LoadH5Seurat(paste0(sp_data_path, 'sp_for_seurat_selected.h5seurat'))
```

```{r}
# the column for annotations
#sc_seu@meta.data$combined = factor(as.character(sc_seu@meta.data$combined))
```

## running Seurat PCA

```{r seurat}
library(dplyr)

sc_seu <- SCTransform(sc_seu, verbose = FALSE) %>% RunPCA(verbose = FALSE)
sp_seu <- SCTransform(sp_seu, verbose = FALSE) %>% RunPCA(verbose = FALSE)

start_time <- Sys.time()

anchors <- FindTransferAnchors(reference = sc_seu, query = sp_seu, normalization.method = "SCT")

predictions.assay <- TransferData(anchorset = anchors, refdata = sc_seu$combined,
                                  weight.reduction = 'pcaproject')

end_time <- Sys.time()
end_time - start_time
```

```{r}
saveRDS(object = predictions.assay,
        file = paste0(results_folder, 'seurat/results_selected_133clusters_pca.RDS'))

predictions.assay$spot = rownames(predictions.assay)
write.csv(predictions.assay, paste0(results_folder, 'seurat/results_selected_133clusters_pca.csv'))
predictions.assay[1:5, 1:5]
```

## running Seurat CCA

```{r seurat_cca}
library(dplyr)

sc_seu <- SCTransform(sc_seu, verbose = FALSE) %>% RunPCA(verbose = FALSE)
sp_seu <- SCTransform(sp_seu, verbose = FALSE) %>% RunPCA(verbose = FALSE)

start_time <- Sys.time()

anchors <- FindTransferAnchors(reference = sc_seu, query = sp_seu, normalization.method = "SCT", reduction='cca')

predictions.assay <- TransferData(anchorset = anchors, refdata = sc_seu$combined,
                                  weight.reduction = 'cca')

end_time <- Sys.time()
end_time - start_time
```

```{r}
saveRDS(object = predictions.assay,
        file = paste0(results_folder, 'seurat/results_selected_133clusters_cca.RDS'))

predictions.assay$spot = rownames(predictions.assay)
write.csv(predictions.assay, paste0(results_folder, 'seurat/results_selected_133clusters_cca.csv'))
predictions.assay[1:5, 1:5]
```
